High-Dimensional Regularized Discriminant Analysis

### 2016-01-02 reports/paper-hdrda.Rnw
Summarized contamination sim results. Finally.

<<knitr_setup, include=FALSE, echo=FALSE, cache=FALSE>>=
# Round digits throughout document including \Sexpr statements.
# See Yihui's response here:
# http://stackoverflow.com/questions/11062497/how-to-avoid-using-round-in-every-sexpr
options(digits = 3)

@@

@@

In this section, we compare our proposed classifier with four classifiers
recently proposed for small-sample, high-dimensional data along with the random
forest from \cite{Breiman:2001fb} using version 3.2.2 of the open-source
statistical software {\tt R}. Within our study, we included penalized linear
discriminant analysis from \cite{Witten:2011kc}, implemented in the {\tt
  penalizedLDA} package. We also considered shrunken centroids regularized
discriminant analysis from \cite{Guo:2007te} in the {\tt rda} package. Because
the {\tt rda} package does not perform the authors' ``Min-Min'' rule
automatically, we applied this rule within our {\tt R} code. We included two
modifications of diagonal linear discriminant analysis from \cite{Tong:2012hw}
and \cite{Pang:2009ik}, where the former employs an improved mean estimator and
the latter utilizes an improved variance estimator. Both classifiers are
available in the {\tt sparsediscrim} package. Finally, we incorporated the
random forest as a benchmark based on the findings of
\cite{FernandezDelgado:2014ul}, who concluded that the random forest is often
superior to other classifiers in benchmark studies. We used the implementation
of the random-forest classifier from the {\tt randomForest} package with 250
trees and 100 maximum nodes. For each classifer we explicitly set prior
probabilities as equal, if applicable. All other classifier options were set to
their default settings. Below, we refer to each classifier by the first
author's surname.

@@

From each of the $K=3$ populations, we sampled 25 training observations ($n_k =
25$ for all $k$) and 10,000 test observations. After training each classifier on
the training data, we classified the test data sets and computed the proportion
of mislabeled test observations to estimate the classification error rate for
each classifier. Repeating this process 500 times, we computed the average of
the error-rate estimates for each classifier. We allowed the number of features
to vary from $p = 100$ to $p = 500$ in increments of 100 to examine the
classification accuracy as the feature dimension increased while maintaining a
small sample size. Furthermore, although \cite{Guo:2007te} consider $\rho_k =
0.9$ for all $k$, we considered $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$
to explore the more realistic assumption of unequal covariance matrices.

@@

As expected, regardless of the values of $p$, $\rho$, or $\bm \mu$, as the
contaminated probability increases, the average error rate increases.  Also, as
can be expected, the average error rates were typically higher for $\bm \mu=0.5$
than for $\bm \mu=1$. The difference in the average error rates for the methods
of Pang, Tong, and Witten all yielded increasingly higher average error rates
than HDRDA ridge, HDRDA complex, Guo, or random forests as the contaminated
probability increased. The correlation coefficient appears to have had little,
if any, effect on the average error rates for any of the considered classifiers.

\begin{figure}
<<sim_results, fig.width=8, fig.height=8, dpi=200>>=
p <- ggplot(results_contaminated, aes(x=contamination_prob, y=Error))
p <- p + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
p <- p + facet_wrap(~ p, labeller = "label_both")
p <- p + xlab("Contamination Probability") + ylab("Error Rate")
p + theme_bw()
@
\caption{Averages of the 1000 error-rate estimates of the competing classifiers
  for each data set as a function of $d$. Approximate standard errors are no
  greater than \Sexpr{max_se}.}
\label{fig:sim-results}
\end{figure}
Regardless of the value of $rho$, $p$, or $\bm \mu$, random forests consistently
yielded the highest average error rate, and the average error rate for random
forests was much more pronounced when $\bm \mu=0.5$. However, as the
contaminated probability increased, random forests became increasingly
competitive, and, in some cases, e.g, $\bm \mu=0.5$, $p=100$, and contaminated
probability greater than 0.05 yielded the lowest average error rate.

Pang and Tong's methods yielded practically the same and consistently the worst
average error rates when the contaminated probability was greater than 0,
regardless of the value of $\bm \mu$, $p$, or $\rho$. Witten's method performed
only slightly better as the contaminated probability increased, however, when
$\bm \mu = 0.5$, Witten's method became increasingly competitive with Pang and
Tong for yielding the highest average error rate as $p$ increased.

Guo's method always outperformed Pang, Witten, and Tong, but never outperformed
the HDRDA methods. For the considered parameter configurations, Guo's method
performed similarly to random forests when the contaminated probability was
greater than 0 and yielded similar results to HDRDA ridge, HDRDA convex when the
contaminated probability was equal to 0. However, as the contaminated
probability increased, Guo's method was often outperformed by HDRDA ridge, HDRDA
convex, and random forests.

HDRDA ridge and HDRDA complex yielded the lowest average error rates, regardless
of the value of $\bm \mu$, $p$, or $\rho$, with HDRDA ridge typically
outperforming HDRDA complex when the contaminated probability was greater than
0. HDRDA ridge only seemed to have competition with random forests when the
contaminated probability was 0.20. Moreover, neither HDRDA ridge nor HDRDA
complex seem to be affected by $p$.

>>>>

<<knitr_setup, include=FALSE, echo=FALSE, cache=FALSE>>=
# Round digits throughout document including \Sexpr statements.
# See Yihui's response here:
# http://stackoverflow.com/questions/11062497/how-to-avoid-using-round-in-every-sexpr
options(digits=3)

@@

# Remove underscore from classifier names
levels(results_contaminated$Classifier) <- gsub("_", " ",
                                                levels(results_contaminated$Classifier))

# Maximum Standard Error for Contaminated Sims
max_sim_se <- with(results_contaminated, max(sqrt(Error * (1 - Error) / 500)))
max_sim_se <- round(max_sim_se, digits=3)

@@

In this section, we compare our proposed classifier with four classifiers
recently proposed for small-sample, high-dimensional data along with the
random-forest classifier from \cite{Breiman:2001fb} using version 3.2.2 of the
open-source statistical software {\tt R}. Within our study, we included
penalized linear discriminant analysis from \cite{Witten:2011kc}, implemented in
the {\tt penalizedLDA} package. We also considered shrunken centroids
regularized discriminant analysis from \cite{Guo:2007te} in the {\tt rda}
package. Because the {\tt rda} package does not perform the authors' ``Min-Min''
rule automatically, we applied this rule within our {\tt R} code. We included
two modifications of diagonal linear discriminant analysis from
\cite{Tong:2012hw} and \cite{Pang:2009ik}, where the former employs an improved
mean estimator and the latter utilizes an improved variance estimator. Both
classifiers are available in the {\tt sparsediscrim} package. Finally, we
incorporated the random forest as a benchmark based on the findings of
\cite{FernandezDelgado:2014ul}, who concluded that the random forest is often
superior to other classifiers in benchmark studies. We used the implementation
of the random-forest classifier from the {\tt randomForest} package with 250
trees and 100 maximum nodes. For each classifer we explicitly set prior
probabilities as equal, if applicable. All other classifier options were set to
their default settings. Below, we refer to each classifier by the first author's
surname. All simulations were conducted on an Amazon Elastic Compute Cloud (EC2)
instance of type `c4.4xlarge`.

@@

From each of the $K=3$ populations, we sampled 25 training observations ($n_k =
25$ for all $k$) and 10,000 test observations. After training each classifier on
the training data, we classified the test data sets and computed the proportion
of mislabeled test observations to estimate the classification error rate for
each classifier. Repeating this process 500 times, we computed the average of
the error-rate estimates for each classifier. We allowed the number of features
to vary from $p = 100$ to $p = 500$ in increments of 100 to examine the
classification accuracy as the feature dimension increased while maintaining a
small sample size. \cite{Guo:2007te} originally considered $\rho_k = 0.9$ for
all $k$. To explore the more realistic assumption of unequal covariance
matrices, we instead put $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$

@@

In Figure \ref{fig:sim-results}, we observe each classifier's average
classification error rates for the values of $\epsilon$ and $p$. Unsurprisingly,
the average error rate increased for each classifier as the contamination
probability $\epsilon$ increased regardless of the value of $p$. Sensitivity to
the presence of outliers was most apparent for the for the Pang, Tong, and
Witten classifiers. For smaller dimensions, the random-forest and HDRDA
classifiers tended to outperform the remaining classifiers with the random
forest performing best. As the feature dimension increased with $p \ge 300$,
both HDRDA classifiers outperformed all other classifiers, suggesting that their
dimension reduction inherent better captured the classificatory information in
the small samples even in the presence of outliers.

\begin{figure}
<<sim_results, fig.width=8, fig.height=8, dpi=200>>=
p <- ggplot(results_contaminated, aes(x=contamination_prob, y=Error))
p <- p + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
p <- p + facet_wrap(~ p, labeller = "label_both")
p <- p + xlab("Contamination Probability") + ylab("Expected Error Rate")
p + theme_bw()
@
\caption{Average classification error rates as a function of the contamination
  probability $\epsilon$. Approximate standard errors were no greater than
  \Sexpr{max_sim_se}.}
\label{fig:sim-results}
\end{figure}

The Pang, Tong, and Witten methods yielded practically the same and consistently
the worst error rates when outliers were present with $\epsilon > 0$, suggesting
that these classifiers were sensitive to outliers. Notice, for example, that
when $p=400$, the error rates of the Pang, Tong, and Witten classifiers
increased dramatically from approximately 19\% when no outliers were present to
approximately 43\% when $\epsilon=0.05$. The sharp increase in average error
rates for these three classifiers continued as $\epsilon$ increased
further. Guo's method always outperformed Pang, Witten, and Tong, but once
outliers were present, the Guo classifier's average error rate was competitive
with neither the HDRDA classifiers nor the random forest.

\begin{figure}
<<sim_results_by_p, fig.width=8, fig.height=8, dpi=200>>=
p2 <- results_contaminated %>%
  mutate(p=as.numeric(as.character(p)),
         contamination_prob=round(contamination_prob, digits=2)) %>%
    filter(contamination_prob %in% c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5)) %>%
    rename(`Contamination Probability`=contamination_prob) %>%
      ggplot(., aes(x=p, y=Error))
p2 <- p2 + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
p2 <- p2 + facet_wrap(~ `Contamination Probability`, labeller="label_both")
p2 <- p2 + xlab("Number of Features") + ylab("Expected Error Rate")
p2 + theme_bw()
@
\caption{Average classification error rates as a function of the number of
  features $p$. Approximate standard errors were no greater than
  \Sexpr{max_sim_se}.}
\label{fig:sim-results-by-p}
\end{figure}

In Figure \ref{fig:sim-results-by-p} we again examine the simulation results as
a function of the number of features $p$ for a subset of the values of
$\epsilon$. This set of plots allows us to investigate the effect of feature
dimensionality on classification performance. When no outliers were present
(i.e., $\epsilon=0$), the random-forest classifier was outperformed by all other
classifiers. Furthermore, the HDRDA classifiers were superior in terms of
average error rate in this setting. As $p$ increased, an elevation in average
error rate was expected for all classifiers, but the increase was not observed
to be substantial.

For $\epsilon > 0$, we observed a different behavior in classification
performance. First, the Pang, Tong, and Witten methods along with the random
forest increased in average error rate as $p$ increased. Contrarily, the
performance of the HDRDA and Guo classifiers was hardly affected by $p$.  Also,
as discussed above, the HDRDA classifiers were superior to all other classifiers
for larger values of $p$ with only the random forest outperforming them for
smaller feature dimensions.
